
\msection{Results from Prior NSF Support}

\textbf{Nick Turk-Browne} was previously supported as co-PI under NSF
grant BCS-1229597, ``MRI: Acquisition of High Performance Compute
Cluster for Multivariate Real-time and Whole-brain Correlation
Analysis of fMRI Data'' from Aug 15, 2012 to July 31, 2015. The PI
of the grant was Jonathan Cohen and the co-PIs were Ray Lee, Kai Li,
Kenneth Norman, and Turk-Browne (Princeton University at the
time). The total award amount was \$527,978.
{\bf Intellectual Merit.} This project resulted in the acquisition of
a high-performance cluster for the analysis of fMRI data.
The cluster has 50 nodes, each with 2 Xeon chips
(16 cores) as host CPUs and 2 Xeon Phi boards (120 cores) as compute nodes (6,800 cores total), as well as 256GB DRAM, 1.6TB Flash memory, and
30TB of local storage.
The Xeon and Phi processors were donated by Intel as
in-kind cost sharing valued at a similar amount. 
This system was integrated with state-of-the-art fMRI, enabling two
exciting long-term projects that required rapid-throughput analysis of
high-dimensionality data: closed-loop real-time neurofeedback~\citep{deBettencourt:2015} and
whole-brain functional connectivity analysis~\citep{Wang:2015}.
{\bf Broader Impact.} This work brought together a diverse team of
neuroscientists, psychologists, physicists, and computer scientists,
enhancing the computing and imaging infrastructure at
Princeton, and its intellectual and educational environment. Many
graduate students, postdoctoral fellows, and faculty have used this
cluster for neuroscience research, including the neuroscience and
computer science graduate students deBettencourt and Wang, who led
the two projects mentioned above. These projects also contributed to the
development of BrainIAK (brainiak.org), an open-source software package
for advanced fMRI analysis, part of our industry partnership with Intel Labs.

%\vskip10pt \noindent Turk-Browne was previously Co-PI of NSF grant
%ACI-1440750, ``CC*IIE Engineer: A Software-Defined Campus Network for
%Big-Data Sciences'' from September 1, 2014 to August 31, 2016. The PI
%of the grant was Jennifer Rexford and the co-PIs were Curt Hillegas,
%Christopher Tully, and Turk-Browne (all Princeton University at the
%time). The total award amount was \$399,776.
%
%{\bf Intellectual Merit.} This project resulted in the hiring of a
%full-time cyber infrastructure engineer (CIE), who has been leading a
%major campus initiative to deploy software-defined networking (SDN)
%for the next generation of data-driven scientific research, including
%in neuroscience, physics, geosciences, and genomics. This initiative
%led to cost-effective scaling of the campus IPS, innovative solutions
%to server load balancing, better monitoring of network performance
%with perSONAR, and testing of OpenFlow switches.
%
%{\bf Broader Impact.} One of the first tasks of the CIE was to create
%a survey that interviewed faculty about their networking needs. This
%survey and subsequent activities under this project led to a unique
%collaboration between researchers in Computer Science and the
%administrative Office of Information Technology, helping ensure that
%academic research needs across campus were central to IT
%infrastructure decisions.

%\vskip10pt \textbf{John Lafferty} was previously supported as co-PI under NSF
%grant IIS-1116730, ``III: Small: Nonparametric Structure Learning for
%Complex Scientific Datasets,'' from August 1, 2011 to July 31, 2014.
%The PI of the grant was Han Liu (Princeton University) and the co-PIs
%were Lafferty and Larry Wasserman (Carnegie Mellon University).  The
%total award amount was \$499,344; the amount awarded to the University
%of Chicago was \$118,750.
%
%{\bf Intellectual Merit.} %The research supported under this grant
%%contributed to the subfield of statistics called nonparametric
%%sparsity, which exploits novel sparsity-inducing regularizations to
%%fit nonparametric models. The project focused on developing scalable
%methods for finding structure in complex scientific datasets, without
%making strong distributional assumptions. The project explored several
%aspects of nonparametric structure learning, including methods,
%theory, large-scale computing, and applications, with five concrete
%aims: (1) nonparametric structure learning in high dimensions, (2)
%nonparametric conditional structure learning, (3) regularization
%parameter selection, (4) parallel and online nonparametric learning,
%and (5) minimax theory for nonparametric structure learning problems. 
%The outcomes included practical models and algorithms; application
%areas included genomics, cognitive neuroscience, climate science,
%astrophysics, and language processing.  Publications resulting from
%this grant include
%\citep{fcca,mehddtgm,chen:13,challenges,quadro,bigp,admm,NRRR:NIPS,gu:12,
%ospif,hdssipca,tpca,pcangd,rspcr,tmehdv,scale,pca,bernoulli,direct,
%coda,spcahdmts,peace,cna,biglasso,fshdc,kolarhan,gemad,gemd,
%lafferty2012,flare,hdtgm,liu2012,skeptic,expoconc,insensitive,mrc,
%eigens,calibrate,blossom,Mishra:2015,patterns,mngm,fclime,joint,retm,
%tests,shender:13,soft,latenttree,tar,sconvex,xu:14,csc,sicec,
%calibratedp,huge,hdngevspnp,semirank,amrbcdm,qnegsm,Zhu:18}.
%
%{\bf Broader Impact.} The broader impact of the project included
%interdisciplinary training for graduate students from biostatistics,
%computer science, statistics, and medical schools, strengthening the
%collaboration and interdisciplinary infrastructure between Carnegie
%Mellon, Johns Hopkins, the University of Chicago and Princeton, and
%broadly disseminating the results from this research in journals from
%all relevant fields. The research had impact outside of machine
%learning and statistics. In a genomic study, PI Liu applied structured
%nonparametric methods to analyze high dimensional genomic data,
%identifying several gene mutations associated with autism.  These
%results were published in Nature \citep{patterns}, and reported in the
%New York Times. In another neuroscience study, the PI developed an
%effective algorithm for predicting Attention Deficit Hyperactive
%Disorder (ADHD) disease.  The research led to several statistical
%software packages in R, including \citep{huge,fclime,flare}, all of
%which are freely available on CRAN.

\vskip10pt \textbf{John Lafferty} is currently supported as PI under NSF
grant DMS-1513594, ``Constrained Statistical Estimation and Inference:
Theory, Algorithms and Applications,'' from June 29, 2015 to July 1,
2018. The total award amount was \$320,000. After two years of the
project, the remainder of the funds were transferred from the
University of Chicago to Yale University, where the PI moved in July
2017.
{\bf Intellectual Merit.}  The project is studying constraints that
are present in complex scientific data analysis problems, but that
have not been thoroughly studied in traditional approaches. Different
aspects of theory, algorithms, and applications of statistical
procedures, with constraints imposed on the storage, runtime, shape,
energy or physics of the estimators and applications. The overall goal
is to develop theory and tools that can help
scientists conduct more effective data analysis. Publications under
this grant include: \citep{ChatterjeeL18,MishraILH18,
abs-1803-01302,MishraLH17,YangB0L16,ChatterjeeDLZ16,ZhengL16,
MishraZLH15,ZhengL15,ZhuL14,Bonak18}.
{\bf Broader Impact.} The impact of this project is aimed in
three directions. First is the development of flexible and principled
large scale data analysis tools that can benefit many scientific
domains. Second, is the development of software that is widely
distributed, allowing others to build on the work. The third is
education, to allow the research to impact the training of students at
both the graduate and undergraduate levels. (Lafferty was previously PI of NSF grant
DMS-1547396, ``RTG: Computational and Applied Mathematics in
Statistical Science'' from July 1, 2016 to July 1, 2017. This grant
did not transfer to Yale University; the current PI is Jonathan Weare
at the University of Chicago. The total award amount is \$1,697,320.)

%{\bf Intellectual Merit.}  This Research Training Group (RTG) project
%supports creation of a dynamic, interactive, and vertically integrated
%community of students and researchers working together in
%computational and applied mathematics and statistics. The work is
%motivated by the growing need to train the next generation of
%statisticians and computational and applied mathematicians in new
%ways, to confront data-centric problems in the natural and social
%sciences.
%
%{\bf Broader Impact.} The broader impact includes vertical integration
%of education and training from undergraduate to postdoctoral
%researchers, including activities at Toyota Technological Institute at
%Chicago and Argonne National Laboratory. Participants in the RTG will
%receive an educational experience that provides them with strong
%preparation for positions in industry, government, and academics, with
%an ability to adopt approaches to problem solving that are drawn from
%across the computational, mathematical, and statistical sciences.

\vskip10pt \textbf{Damon Clark} is currently supported as PI under NSF
grant IOS-1558103, ''Understanding how Neural Nonlinearities Tune
Motion Detection in the Fly Eye,'' from July 1, 2016 to June 30, 2019.
The total award amount is \$461,262.
{\bf Intellectual Merit.} This project focuses on the nonlinearity
in visual motion detection circuits of fruit flies. We are characterizing
the nonlinearity with behavioral and neural
measures, dissecting how individual neurons in the circuit contribute,
and investigating how it relates to the
performance of the circuit with naturalistic inputs. We have so far
published a review~\citep{clark:16} and a methods
paper~\citep{mano:17}, with three more manuscripts
submitted or under review.
{\bf Broader Impact.} Three high school students worked in the lab
over the summers (2016-2018), and completed lessons in Yale's
\textit{Pathways to Science} program focused on what visual
illusions teach us about the brain.

\vskip10pt \textbf{Jeffrey Brock} is currently supported as PI under NSF
grant DMS-1608759, ''Rigidity, Volume, and Combinatorics in Hyperbolic Geometry,'' from Sept 1, 2016 to Aug 31, 2019. The total award amount is \$239,997.
\textbf{Intellectual Merit.} 
Research on deformation theory of hyperbolic 3-manifolds, renormalized volume, and Teichm\"uller geometry appeared in \citet{BMNS:bounded:models,Brock:Bromberg:vol,Brock:Bromberg:cone:inflex,BLMR:wp:limit,BLMR:teichmuller, Brock:Dunfield:norms}. Results included sufficient conditions for bounded geometry, a new direct relation of Weil-Petersson distance and volume, complex limits of rays in Teichm\"uller space, and comparison of Thurston and harmonic norms on the cohomology of hyperbolic 3-manifolds. \citet{Keshavamurthy:TBI} adapted machine learning methods to auto detect CT-identifiable TBI lesions.
\textbf{Broader Impact.} PI served as inaugural director of Brown's DSI, launching its Master's program, coordinating campus outreach, and speaking on alumni panels in San Francisco and Providence. PI was Co-PI and Associate Director of ICERM and will continue as Scientific Advisory Board member. PI was lead organizer of Summer\verb+@+ICERM undergraduate program on topological data analysis and sponsored DSI undergraduate \textit{Datathon} (spring 2018), engaging 200 students from Brown and beyond in an overnight hackathon. 
