\msection{Relation to TRIPODS Project} 

Brown's TRIPODS award, \emph{Foundations of Model Driven Discovery from Massive Data}, focuses on the interplay between the analysis of data arising from a coherent data generating model or mechanism and the refining of such models from tools in data analysis. The focus areas of Brown's TRIPODS activities include causal inference, graphs and networks, and geometry and topology of data. Each of these areas fits thematically with the proposed project and its objectives.

Causal inference is directly involved in deriving neuronal causation of behaviors or actions, and its geometric and topological study figures directly in attempts to relate neural network topology to this causal structure.
Likewise, inferring the structure of graphs and networks form restricted samples is a central challenge of relating fMRI or other neural imaging data to neural structure. 
Finally, methods from Topological Data Analysis have been applied fruitfully to understand the structure and function of place cells in the hippocampus, and relating their topology and geometry to spacial external geometry is an important ongoing area of research.

The activities supported by the TRIPODS award include intensive visits, Seminars, and Summer or Winter Schools, each of which will naturally support the theme of the proposed project, as a ``research domain'' to which methods in each TRIPODS focus area can fruitfully be applied. The proposed activities involve visits between teams of researchers and their students at Yale and at Brown, naturally dovetailing with proposed activities in the TRIPODS project. 

Finally, the TRIPODS institute at Brown is housed within its Data Science Initiative. This Initiative co-sponsors postdoctoral fellows with the TRIPODS Institute, in addition to hiring new faculty in core departments of Applied Mathematics, Biostatistics, Computer Science and Mathematics. The TRIPODS+X workshop at Brown will be housed within the Data Science Initiative, which will be co-located with the new Carney Institute for Brain Science at Brown, in the new site above the Bookstore at 425 Angell Street. This co-location will provide natural linkages between TRIPODS faculty and postdocs and faculty in Brain Science at Brown, providing a rich atmosphere for TRIPODS and TRIPODS+X activities focused on connections between Data Science and Neuroscience.
