\documentclass[pdftex,12pt]{article}

\usepackage{fullpage,amssymb,amsmath}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{times}
\usepackage{enumerate}
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage[pdftex,hyperfootnotes=false]{hyperref}
\hypersetup{pageanchor=false}
\usepackage{marvosym,stmaryrd}

\let\hat\widehat
\let\tilde\widetilde
\renewcommand{\P}{\mbox{$\mathbb{P}$}}
\newcommand{\reals}{\mbox{$\mathbb{R}$}}
\newcommand{\E}{\mbox{$\mathbb{E}$}}
\newcommand{\V}{\mbox{$\mathbb{V}$}}

\bibpunct{(}{)}{;}{a}{,}{,}
\bibliographystyle{ims}

\newcounter{proposal}
\newcounter{project}
\newcounter{subproposal}
\newcounter{subsubproposal}

\def\section#1{\vskip15pt\stepcounter{proposal}\setcounter{subproposal}{0}\noindent
   {\bf\normalsize \arabic{proposal}.\enspace #1}\vskip1pt}

\renewcommand{\baselinestretch}{1.0}

\parskip1pt
\parindent0pt

\begin{document}

\parindent=0pt
\pagestyle{empty}
\vskip-5pt
\noindent {\large\sc John D. Lafferty -- Biographical sketch}
\def\surl#1{{\small\sl\url{#1}}}
\vskip2pt{\bf Address}

\begin{quote}
Department of Statistics and Data Science\\
Yale University\\
24 Hillhouse Avenue\\
New Haven, CT 06511\\
Email: {john.lafferty@yale.edu}
\end{quote}


\vskip1pt{\bf Professional Preparation}

\begin{quote}


1982--1986, Princeton University\\
Ph.D. in Mathematics, 1986,  M.A. in Mathematics, 1984\\
Member of the Program in Applied and Computational Mathematics\\
Thesis advisor: Edward Nelson
\vskip3pt

1981--1982, Massachusetts Institute of Technology\\
Undergraduate work in mathematics and statistics.  Beinecke Fellowship.
\vskip3pt

1978--1981, Middlebury College \\
B.A. 1982 {\em summa cum laude\/} with highest
honors in Mathematics.
\vskip5pt

\end{quote}


\vskip1pt{\bf Appointments}

\begin{quote}
2017--Present, Yale University \\
John C. Malone Professor of Statistics and Data Science\\
Secondary appointment, Department of Computer Science
\vskip5pt

2013--Present, Toyota Technological Institute at Chicago \\
Adjoint Professor
\vskip5pt

2011--2017, University of Chicago \\
Louis Block Professor of Statistics and Computer Science
\vskip5pt

1994--2011,  Carnegie Mellon University\\
Professor of Computer Science, Machine Learning, and Statistics\\[1pt]
2005--2008, Co-Director, Ph.D. Program in Machine Learning\\
1998--2004, Associate Professor of Computer Science\\
1994--1997, Assistant Research Professor
\vskip5pt

1988--1994, IBM Thomas J. Watson Research Center,
Yorktown Heights, NY\\
Research Staff Member, Department of Computer Sciences
\vskip5pt

1986--1987, Harvard University\\
Assistant Professor, Benjamin Pierce Lecturer on Mathematics
\end{quote}


\def\secskip{\vskip6pt}

\secskip{\bf Publications Related to Proposal}
\begin{enumerate}

\item Y. Zhu  and J. Lafferty, ``Quantized minimax estimation over Sobolev ellipsoids,''
Information and Inference: A Journal of the IMA, Volume 7, Issue 1, 15
March 2018, pp 31--82.

\item Y. Zhu  and J. Lafferty, ``Distributed nonparametric regression
  under communication constraints,'' arXiv:1803.01302, to appear in 
International Conference on Machine Learning (ICML), 2018.

\item N.~Mishra, H.~Zhang, J.~Lafferty and H.~Hoffman, ``A graphical
  model approach to  minimizing energy under performance constraints,''
  Architecural Support for Programming Languages and Operating
  Systems (ASPLOS), 2015.

\item D.~Blei and J.~Lafferty,
``Dynamic topic models,''
 {\it Machine Learning: Proceedings of the Twenty-Third International
  Conference (ICML)}, pp.~113--120, 2006.

\item X. Zhu, Z. Ghahramani and J. Lafferty, ``Semi-supervised
  learning using Gaussian fields and harmonic functions,'' {\it
    Proceedings of the International
    Conference on Machine Learning}, pp. 912--919, 2003.
\end{enumerate}


\secskip{\bf Other Publications}
\begin{enumerate}
\item M.~Xu, M.~Chen and J.~Lafferty, 
``Faithful variable selection for high dimensional convex
  regression,'' 
{\it Ann. Statist.}, Vol.~44, Number 6 (2016), 2624--2660.

\item H. Liu, F. Han, M. Yuan, J. Lafferty, and L. Wasserman,
``High-dimensional semiparametric Gaussian copula graphical models'',
{\it Ann. Statist.}, Vol.~40, Number 4 (2012), 2293--2326.

\item P.~Ravikumar, M.~Wainwright and J.~Lafferty, 
 ``High dimensional Ising model selection using $\ell_1$ regularized
 logistic regression,'' 
 {\it Anna.~Statist.} Volume 38, Number 3 (2010), 1287--1319.

\item P.~Ravikumar, J.~Lafferty, H.~Liu, and L.~Wasserman,
``Sparse additive models,'' {\it Journal of the Royal
  Statistical Society: Series B}, 
  Volume 71, part 5, pages 1009--1030, 2009.

\item
S.~Zhou, J.~Lafferty and L.~Wasserman, 
``Compressed and privacy sensitive sparse regression,''
{\it IEEE Trans. Information Theory}, Vol.~55, Number 2, February
2009, pages 846--866.
\end{enumerate}



\secskip{\bf Synergistic Activities}
\begingroup
\addtolength{\leftmargini}{-.1in}
\begin{quote}
Member of the board of the NIPS Foundation, 2010--present.
Member of the Committee on Applied and Theoretical
Statistics (CATS), National Research Council of the National Academy
of Sciences, 2011--2016.
Member of NORC Advisory Committee on Statistics, Machine Learning and
High Performance Computing, 2014--present.
\end{quote}
\endgroup

\secskip{\bf Recent Collaborators and Other Affiliations}

\begingroup
\addtolength{\leftmargini}{-.1in}
\begin{quote}
David Blei (Princeton),  Rina Barber (Univ.~Chicago), Sabyasachi
Chatterjee (Univ.~Illinois), John Duchi (Stanford), Chao Gao
(Univ.~Chicago), Han Liu (Northwestern),
Larry Wasserman (CMU), Min Xu (Rutgers), Yuancheng Zhu (Univ.~Pennsylvania).
\end{quote}
\endgroup

\end{document}
